---
title: "Werba_Zooplankton Figures"
author: "Jo"
date: "June 28, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dotwhisker)
library(grid)
library(gridExtra)
source_rmd <- function(file, local = FALSE, ...){
  options(knitr.duplicate.label = 'allow')

  tempR <- tempfile(tmpdir = ".", fileext = ".R")
  on.exit(unlink(tempR))
  knitr::purl(file, output=tempR, quiet = TRUE)

  envir <- globalenv()
  source(tempR, local = envir, ...)
}
source_rmd("WerbaMasterCleanCode.Rmd")
load("decomposition2.Rdata")
```

Here I will provide all the code for the graphs. 


Theme set
all graphs (except decomposition) only show results for the mixed dispersal treatments so that the graphs can be in 2 dimensions. 

```{r}
theme_set(theme_bw()) 
theme_update(axis.text.x = element_text(size = 12),
             axis.text.y = element_text(size = 12),
             axis.title.x = element_text(size = 14),
             axis.title.y = element_text(size = 14),
             legend.title = element_text(size = 10),
             legend.text = element_text(size = 8),
             legend.spacing = unit(0.25, "cm"),
             legend.key.width = unit(0.35, "cm"),
             panel.grid.major = element_blank(),
             panel.grid.minor = element_blank(),
             strip.background = element_blank(),
             panel.spacing = unit(0, "lines"),
             #      legend.key.size = unit(.15, "cm"),
             legend.key = element_rect(fill = "white"),
             panel.spacing.y = unit(-0.75, "lines"),
             panel.border = element_rect(colour = "black", 
                                         fill = NA, size = 1),
             strip.text.x = element_text(size = 18, colour = "black", 
                                         face = "bold"))

mycolors<-c("dodgerblue4","darkorange","darkorchid1","red")
mylines<-c("solid","dashed","dotdash","dotted")
names(mycolors) <- levels(bigdata$Salinity)
colScale <- scale_colour_manual(name = "Salinity",values = mycolors)
ColFill<-scale_fill_manual(name = "Salinity",values = mycolors)
ltype<-scale_linetype_manual(name="Salinity", values=mylines)
```


Effective Diversity
tsummX (where X is the salinity treatment) has the averages and sd of the measured values

tdata2 has the predicted values from the model and the error around the model
```{r}
g0<- ggplot(tdata2, aes(Date2,true)) 
g1<-g0+geom_point(data=tsumm0, aes(y=true,x=Date2),colour="dodgerblue4",size=1.25)+
  geom_errorbar(data = tsumm0,aes(x=Date2,ymin=true-sd/2,ymax=true+sd/2),colour="dodgerblue4",
                size=0.5, width=1)+
  geom_point(data=tsumm5,aes(y=true,x=Date2),colour="darkorange",size=1.25)+
  geom_errorbar(data = tsumm5,aes(x=Date2,ymin=true-sd/2,ymax=true+sd/2),
                colour="darkorange",size=0.5,width=1)+
  geom_point(data=tsumm9,aes(y=true,x=Date2),colour="darkorchid1",size=1.25)+
  geom_errorbar(data = tsumm9,aes(x=Date2,ymin=true-sd/2,ymax=true+sd/2),colour="darkorchid1",size=0.5, width=1)+
  geom_point(data=tsumm13, aes(y=true,x=Date2),colour="red",size=1.25)+geom_errorbar(data = tsumm13,aes(x=Date2,ymin=true-sd/2,ymax=true+sd/2),colour="red",size=0.5, width=1)

g2<-g1+ geom_ribbon(data=tdata2,aes(ymin = LL, ymax = UL, fill = Salinity), alpha = .1) +
  geom_line(data=tdata2,aes(colour = Salinity,linetype=Salinity), size = 2) +
  labs(x = "Date2", y = "True")+colScale+ColFill+ltype

eff <- g2+ylab("Diversity")+xlab(" ")+
          theme(legend.position = c(0.75,0.75))+
    scale_colour_manual(guide= guide_legend(direction = "horizontal", title.position = "top"),
                          values = mycolors) +
    scale_linetype_manual(guide= guide_legend(direction = "horizontal"),
                            values = mylines) +
    scale_fill_manual(guide= guide_legend(direction = "horizontal"),
                      values = mycolors)
```
Example of how predicted values and error on the model was found for graphing; this code was used throughout

```{r}
newdat1 <- expand.grid(Dispersal=unique(bigdata$Dispersal),
                      Salinity=unique(bigdata$Salinity),Date2=unique(bigdata$Date2), sreal= 0, true=0)
mm = model.matrix(terms(trudiv),newdat1)
newdat1$true = mm %*% coef(trudiv)
pvar1 <- diag(mm %*% tcrossprod(vcov(trudiv),mm))
newdat1 <- data.frame(
  newdat1
  , plo = newdat1$true-sqrt(pvar1)
  , phi = newdat1$true+sqrt(pvar1)
)

```


Richness
rsummX (where X is the salinity treatment) has the averages and sd of the measured values

rdata2 has the predicted values from the model and the error around the model
```{r}
r<- ggplot(rdata2, aes(Date2,rich)) 
r1<-r+geom_point(data=rsumm0, aes(y=rich,x=Date2),colour="dodgerblue4",size=1.25)+
  geom_errorbar(data = rsumm0,aes(x=Date2,ymin=rich-SE,ymax=rich+SE),colour="dodgerblue4",
                size=0.5, width=1)+
  geom_point(data=rsumm5,aes(y=rich,x=Date2),colour="darkorange",size=1.25)+
  geom_errorbar(data = rsumm5,aes(x=Date2,ymin=rich-SE,ymax=rich+SE),
                colour="darkorange",size=0.5,width=1)+
  geom_point(data=rsumm9,aes(y=rich,x=Date2),colour="darkorchid1",size=1.25)+
  geom_errorbar(data = rsumm9,aes(x=Date2,ymin=rich-SE,ymax=rich+SE),colour="darkorchid1",size=0.5, width=1)+
  geom_point(data=rsumm13, aes(y=rich,x=Date2),colour="red",size=1.25)+
  geom_errorbar(data = rsumm13,aes(x=Date2,ymin=rich-SE,ymax=rich+SE),colour="red",size=0.5, width=1)


r2<-r1+ geom_ribbon(data=rdata2,aes(ymin = LL, ymax = UL, fill = Salinity), alpha = .1) +
  geom_line(data=rdata2,aes(colour = Salinity,linetype=Salinity), size = 0.75) +
  labs(x = "Date2", y = "rich")+colScale+ColFill+ltype

rg<-r2+ylab("Richness")+xlab("Day")+
    theme(legend.position = "none")
    

 
grid.arrange(eff,rg, ncol=1)
```

Abundance
adata2 is the predicted values and error from the model.
I include all data points because the error around averages won't be distributed normally.

```{r}
g0<- ggplot(adata2, aes(Date2,abundance)) +
  geom_ribbon(aes(ymin = LL, ymax = UL, fill = Salinity), alpha = .25) +
  geom_line(aes(colour = Salinity,linetype=Salinity), size = 0.75) +
  labs(x = "Date2", y = "Abundance")+colScale+ColFill+ltype
g1<-g0+geom_jitter(data=bigdata2, aes(y=abundance, x=Date2, color=Salinity),size=1.25, width=2)
g1+ylab("Abundance")+xlab("Day")+theme(legend.position = c(0.5,0.8))+
  scale_colour_manual(guide= guide_legend(direction = "horizontal", title.position = "top"),
                      values = mycolors) +
  scale_linetype_manual(guide= guide_legend(direction = "horizontal"), values = mylines) +
  scale_fill_manual(guide= guide_legend(direction = "horizontal"),
                    values = mycolors)
```


For the supplement and for interpretation ease I made coef plots for each of the models (effective diversity and richness in one plot)

```{r}
d <- dwplot(list(trudiv,RichMod1))+
  relabel_y_axis(c("Dispersal", "Salinity", expression("Salinity"^{2}), 
                     "Date", expression("Date"^{2}), "Salinity:Date", expression(Salinity^{2}:Date), expression(Salinity:Date^{2}), expression(Salinity^{2}:Date^{2}))) +
      xlab("Coefficient Estimate") + ylab("") +
     geom_vline(xintercept = 0, colour = "grey60", linetype = 2) +
     ggtitle("Effective Diversity and Richness") +
     theme(plot.title = element_text(face="bold"),
           legend.justification=c(0.1, 0.9), legend.position=c(0.1, 0.9),
           legend.background = element_rect(colour="grey80"),
           legend.title = element_blank()) + scale_colour_discrete(
                      name = " ",
                      breaks = c("Model 1", "Model 2"),
                      labels = c("Effective Diversity", "Richness"))

print(d)

ab <- dwplot(am1)+
  relabel_y_axis(c("Dispersal", "Salinity", expression("Salinity"^{2}), 
                     "Date", "Salinity:Date", expression(Salinity^{2}:Date))) +
      xlab("Coefficient Estimate") + ylab("") +
     geom_vline(xintercept = 0, colour = "grey60", linetype = 2) +
     ggtitle("Abundance") +
     theme(plot.title = element_text(face="bold"))

print(ab)

```


initial and final communities all treatments-- no graphs print until the end--

```{r}

limits = aes(ymax = agg$mean_mds2 + agg$sd_mds2, ymin=agg$mean_mds2 - agg$sd_mds2) 
limitsx = aes(xmax = agg$mean_mds1 + agg$sd_mds1, xmin=agg$mean_mds1 - agg$sd_mds1)

nc <- ggplot(agg, aes(mean_mds1, mean_mds2, label = TRUE)) +
  geom_point(data = agg, aes(colour = dispersal,pch=salinity), size=5)
nc <- nc+geom_errorbar(limits, width=0.05, size = 0.25)+xlab(" ") +
  ylab("Axis 2")
Initial_comm <- nc+geom_errorbarh(limitsx, height=0.1, size = 0.25) +
  ggtitle("a. Initial") +
    scale_shape_manual(values = c(16,17,15,18))+ 
    theme(legend.justification = c(0.07, 0.9), legend.position = c(0.07,0.9))+
    theme(legend.key.size = unit(.25, "cm"))+
    theme(legend.box = "horizontal")+
    scale_color_discrete(labels=c("Mixed", "Single"))+
  guides(colour = guide_legend(title = "Dispersal",
                               override.aes = list(size=3)))+
  guides(shape = guide_legend(title = "Salinity",
                              override.aes = list(size=3)))


#find limits
limits1=aes(ymax = agg1$mean_mds2 + agg1$sd_mds2, ymin=agg1$mean_mds2 - agg1$sd_mds2) 
limitsx1=aes(xmax = agg1$mean_mds1 + agg1$sd_mds1, xmin=agg1$mean_mds1 - agg1$sd_mds1)



nc1 <- ggplot(agg1, aes(mean_mds1, mean_mds2, label = TRUE)) +
  geom_point(data = agg1, aes(colour = dispersal,pch=as.factor(salinity), size=5))
nc2 <- nc1+geom_errorbar(limits1,width=0.05, size = 0.25)+xlab("Axis 1") + ylab("Axis 2")
Final_comm <- nc2+geom_errorbarh(limitsx1,height=0.1, size= 0.25)+ggtitle("b. Final")+
    scale_shape_manual(values = c(16,17,15,18)) +  theme(legend.position = "none")
```


By treatment
```{r}
# need limits for x,y bars 

frlimx <-with(fresh_3,aes(xmax = mean_axis1 + sd_axis1, xmin=mean_axis1 - sd_axis1))
frlimy <-with(fresh_3,aes(ymax = mean_axis2 + sd_axis2, ymin=mean_axis2 - sd_axis2))

flimx <-with(five_3,aes(xmax = mean_axis1 + sd_axis1, xmin=mean_axis1 - sd_axis1))
flimy <-with(five_3,aes(ymax = mean_axis2 + sd_axis2, ymin=mean_axis2 - sd_axis2))

nlimx <-with(nine_3,aes(xmax = mean_axis1 + sd_axis1, xmin=mean_axis1 - sd_axis1))
nlimy <-with(nine_3,aes(ymax = mean_axis2 + sd_axis2, ymin=mean_axis2 - sd_axis2))

tlimx <-with(teen_3,aes(xmax = mean_axis1 + sd_axis1, xmin=mean_axis1 - sd_axis1))
tlimy <-with(teen_3,aes(ymax = mean_axis2 + sd_axis2, ymin=mean_axis2 - sd_axis2))

#graphing for reals

fr <- ggplot(fresh_3, aes(mean_axis1,mean_axis2, label=TRUE)) +
  geom_point(aes(color=Dispersal, pch= as.factor(Date)), size = 5)

fr1 <- fr + geom_errorbar(frlimy, width=0.1, size = 0.25) +
       geom_errorbarh(frlimx, height=0.1, size = 0.25)
  
fresh_graph <- fr1 + ggtitle("a. Freshwater")+xlab(" ") + ylab("Axis 2") +
    theme(legend.position = "none")



fg <- ggplot(five_3, aes(mean_axis1,mean_axis2, label=TRUE)) +
  geom_point(aes(color=Dispersal, pch= as.factor(Date)), size = 5)

fg1 <- fg + geom_errorbar(flimy, width=0.1, size = 0.25) +
  geom_errorbarh(flimx, height=0.1, size = 0.25)

five_graph <- fg1 + ggtitle("b. Five Salinity")+xlab(" ") + ylab(" ") +
    theme(legend.position = "none")

ng <- ggplot(nine_3, aes(mean_axis1,mean_axis2, label=TRUE)) +
  geom_point(aes(color=Dispersal, pch= as.factor(Date)), size = 5)

ng1 <- ng + geom_errorbar(nlimy, width=0.1, size = 0.25) +
  geom_errorbarh(nlimx, height=0.1, size = 0.25)

nine_graph <- ng1 + ggtitle("c. Nine Salinity")+xlab("Axis 1") + ylab("Axis 2") +
    theme(legend.position = "none")

tg <- ggplot(teen_3, aes(mean_axis1,mean_axis2, label=TRUE)) +
  geom_point(aes(color=Dispersal, shape= as.factor(Date)), size = 5)

tg1 <- tg + geom_errorbar(tlimy, width=0.1, size = 0.25) +
  geom_errorbarh(tlimx, height=0.1, size = 0.25)

teen_graph <- tg1 + ggtitle("d. Thirteen Salinity")+xlab("Axis 1") + ylab(" ") +
    scale_color_discrete(labels=c("Mixed", "Single")) +
    scale_shape_discrete(labels=c("Initial", "Final"))+
    theme(legend.justification = c(0.04, 0.95), legend.position = c(0.04,0.95))+
    theme(legend.box = "horizontal")+
    theme(legend.key.size = unit(0.20, "cm")) +
    guides(colour = guide_legend(override.aes = list(size=3)))+
    guides(shape = guide_legend(title = "Date",
                              override.aes = list(size=3)))
```


Make into two grids
```{r}

grid.arrange(fresh_graph,five_graph, nine_graph, teen_graph, ncol=2)

grid.arrange(Initial_comm, Final_comm, ncol=1)
```



Decomposition


Graph -- newdat is a dataframe with all the predictions from models. rawdata means has the measured means and standard deviations
```{r}
ggplot(newdat, aes(decomp.Salinity, exp(pred))) + geom_line(aes(colour = as.factor(decomp.Dispersal))) + 
  geom_point(data = rawdatmeans, aes(decomp.Salinity, exp(pred), colour=as.factor(decomp.Dispersal))) + 
  geom_errorbar(data = rawdatmeans, aes(decomp.Salinity, ymin=exp(pred -stdv), ymax=exp(pred+stdv),
                                        color=as.factor(decomp.Dispersal)),
                width=0.1, lwd=1)+
  facet_grid(Type~.)+
  theme(panel.margin.x = unit(1.0, "lines"), panel.margin.y = unit(0.15, "lines"),
        strip.text.y = element_text(size = 14, colour = "black", 
                                    face = "bold", angle = 270),
        axis.text.y = element_text(size = 12.5))+
  ylab("Proportion Remaining")+xlab("Salinity")+
  guides(color= guide_legend(title = "Dispersal 
Treatment")) + scale_color_discrete(labels=c("Mixed", "Single"))+
  theme_bw()

```


Here is a coefplot for the decomposition models

```{r}
dc <- dwplot(list(maple, spartina,phrag))+
      xlab("Coefficient Estimate") + ylab("") +
     geom_vline(xintercept = 0, colour = "grey60", linetype = 2) +
     ggtitle("Decomposition") +
     theme(plot.title = element_text(face="bold"),
           legend.justification=c(0.9, 0.9), legend.position=c(0.9, 0.9),
           legend.background = element_rect(colour="grey80"),
           legend.title = element_blank()) + scale_colour_discrete(
                      name = " ",
                      breaks = c("Model 1", "Model 2", "Model 3"),
                      labels = c("Maple", "Spartina", "Phragmites"))

```

Nitrogen
```{r}
g0<-ggplot(data=true_pred, aes(true,nitr))+geom_point(data=nit_summ,aes(true,nitr,colour=Salinity))+
  geom_ribbon(aes(ymin = LL, ymax = UL, fill = Salinity), alpha = .25) + 
  geom_line(aes(colour = Salinity,linetype=Salinity), size = 2)+colScale+ColFill+ltype+facet_wrap(~Date2)
g0+ylab("Nitrogen")+xlab("Effective Diversity")+
  theme(legend.position = c(0.83,0.85))+
  scale_colour_manual(guide= guide_legend(direction = "horizontal", title.position = "top"),
                      values = mycolors) +
  scale_linetype_manual(guide= guide_legend(direction = "horizontal"), values = mylines) +
  scale_fill_manual(guide= guide_legend(direction = "horizontal"),
                    values = mycolors)


```

Nitrogen coef plot
```{r}
nit <- dwplot(nitmod1)+
  relabel_y_axis(c("Dispersal", "Effective Diversity", "Salinity", 
                     "Date","Salinity:Date")) +
      xlab("Coefficient Estimate") + ylab("") +
     geom_vline(xintercept = 0, colour = "grey60", linetype = 2) +
     ggtitle("Nitrogen") +
     theme(plot.title = element_text(face="bold"))
print(nit)
```

Salinity over time
```{r}
s <- ggplot(bigdata,aes(Date2, sreal, color = as.factor(Salinity)))+geom_point()

s1 <- s + xlab("Day")+ ylab("Salinity") + 
  guides(color= guide_legend(title = "Salinity Treatment", title.position = "top")) +
    theme(legend.position = c(0.83, 0.87), legend.direction = "horizontal") +ColFill+colScale

print(s1)
```







