---
title: "Bacterioplankton response to salinity (aka 2015 CSI Dispersal Experiment)"
author: "Alex Stucy, Jo Werba, Mike McCoy, Ariane Peralta"
date: "Last updated on `r format(Sys.time(), '%d %B, %Y')`"
output: 
  pdf_document:
  fig_caption: true
---
Project Description: TBD

FIGURES SCRIPT (stat script = Thesis_Chunk_RMD.Rmd)


Chunk Description: (I) How does salinity and dispersal influence bacterial diversity?
-Use full data set (with source tanks)
-Calculate Shannon's H' (diversity metric)
-Run stat model diversity~dispersal + salinity + dispersal*salinity
-GRAPH DIVERSITY (NOTE: which part of Jo's code corresponds to this? I'm trying to follow it but having a hard time)
-BULLET POINT RESULTS
```{r (I)}
knitr::opts_chunk$set(echo = TRUE)
require("ggplot2")
require("dotwhisker")
require("grid")
require("gridExtra")
source_rmd <- function(file, local = FALSE, ...){
  options(knitr.duplicate.label = 'allow')

  tempR <- tempfile(tmpdir = ".", fileext = ".R")
  on.exit(unlink(tempR))
  knitr::purl(file, output=tempR, quiet = TRUE)

  envir <- globalenv()
  source(tempR, local = envir, ...)
}
source_rmd("Thesis_Chunk_RMD.Rmd") #this does not run -- the error I receive is "Error in setwd("~/GitHub/CSI_Dispersal/analyses") : cannot change working directory"
#load("decomposition2.Rdata") this is in Jo's code but not sure what I should have here in place

```
 Theme set all graphs
 
```{r (I)}
theme_set(theme_bw()) 
theme_update(axis.text.x = element_text(size = 12),
             axis.text.y = element_text(size = 12),
             axis.title.x = element_text(size = 14),
             axis.title.y = element_text(size = 14),
             legend.title = element_text(size = 10),
             legend.text = element_text(size = 8),
             legend.spacing = unit(0.25, "cm"),
             legend.key.width = unit(0.35, "cm"),
             panel.grid.major = element_blank(),
             panel.grid.minor = element_blank(),
             strip.background = element_blank(),
             panel.spacing = unit(0, "lines"),
             #      legend.key.size = unit(.15, "cm"),
             legend.key = element_rect(fill = "white"),
             panel.spacing.y = unit(-0.75, "lines"),
             panel.border = element_rect(colour = "black", 
                                         fill = NA, size = 1),
             strip.text.x = element_text(size = 18, colour = "black", 
                                         face = "bold"))

mycolors<-c("dodgerblue4","darkorange","darkorchid1","red")
mylines<-c("solid","dashed","dotdash","dotted")
#next line did not run
names(mycolors) <- levels(bigdata$Salinity)
#below ran successfully
colScale <- scale_colour_manual(name = "Salinity",values = mycolors)
ColFill<-scale_fill_manual(name = "Salinity",values = mycolors)
ltype<-scale_linetype_manual(name="Salinity", values=mylines)
```
