---
title: "Bacterioplankton response to salinity (aka 2015 CSI Dispersal Experiment)"
author: "Alex Stucy, Jo Werba, Mike McCoy, Ariane Peralta"
date: "Last updated on `r format(Sys.time(), '%d %B, %Y')`"
output: 
  pdf_document:
  fig_caption: true
---
Project Description: 


set working directory - Alex you can make a directory on your computer that has the same path so that we can use the same source code - use bash to mkdir GitHub and then clone repos

```{r}
rm(list = ls())
setwd("~/GitHub/CSI_Dispersal/analyses")
```

set source R tools
```{r}
source("../bin/DiversityFunctions.R")
source("../bin/MothurTools.R")
```

load required packages
```{r}
require("vegan")
require("dplyr")
require("nlme")
require("reshape2")
require("ecodist")
require("ggplot2")
require("ade4")
require("png")
require("MASS")
require("grid")
require("ape")
require("png")
require("picante")
```

set standard error function
```{r}
se <- function(x, ...) {
sd(x, na.rm = TRUE)/sqrt(length(na.omit(x)))
}
ci <- function(x, ...) {
1.96 * sd(x, na.rm = TRUE)
}
```

load design files
```{r}
design <- read.csv("../data/CSI_Design_ENV_NoSourceTanks.csv", row.names=1)
head(design)
str(design)

design_crobes <- read.csv("../data/design_CSI.csv", row.names=1)
design_crobes <- design_crobes[-c(grep("mock community", design_crobes$CSI_ID)), ]
dim(design_crobes)
```

load bacterial otu file
```{r}
csi_otu <- read.otu("../data/CSI.shared")
dim(csi_otu)
```

load files Alex source (but change to match AP source and delete this chunk)
```{r}
design <- read.csv("../data/CSI_Design_ENV_NoSourceTanks.csv",row.names=1)
design

#load design file with microbes 
design_crobes <- read.csv("../data/design_CSI.csv",row.names=1)
design_crobes

#subset without mock community
design_crobes <- design_crobes[-c(grep("mock community", design_crobes$CSI_ID)), ]
dim(design_crobes)
```

Alex source mothur tools and otu file (but change to match AP source and delete this chunk)
```{r}
#set source mothur tools
source("../bin/DiversityFunctions.R")
source("../bin/MothurTools.R")

csi_otu <- read.otu("../data/CSI.shared")
dim(csi_otu)


```

remove extra site in design (CSI041)
```{r}
missing <- setdiff(rownames(design_crobes), rownames(csi_otu))
design_crobes <- design_crobes[-(which(rownames(design_crobes) == missing)), ]
dim(design_crobes)

design <- design[-(which(rownames(design) == missing)), ]
dim(design)
```

id source tanks where number = 1,2,3
```{r}
temp <- rownames(design_crobes[which(design_crobes$Number %in% c(1,2,3)), ])
```

remove samples from tank numbers 1,2,3
```{r}
bac.design <- design_crobes[-(which(rownames(design_crobes) %in% temp)), ]
design_crobes2 <- droplevels(bac.design)
dim(design_crobes2)
```

remove source tanks where number = 1,2,3
```{r}
csi_otu_table <- csi_otu[-(which(rownames(csi_otu) %in% temp)), ]
dim(csi_otu_table)
```

remove otu's w/ < 2 occurrences across all sites
```{r}
otu_removal <- csi_otu_table[, which(colSums(csi_otu_table) >= 2)]
dim(otu_removal)
```

removed low abundance samples
```{r}
csi_low_remov <- otu_removal[which(rowSums(otu_removal) >= 13000), ]
dim(csi_low_remov)
```

odd sites in bacterial composition data (101) and remove 041 in design file
```{r}
odd.sites <- c("CSI101")

otu_final <- csi_low_remov[setdiff(rownames(csi_low_remov), odd.sites), ]
design_final <- design[setdiff(rownames(design), odd.sites), ]

all.equal(rownames(design_final), rownames(otu_final))
```

create presence absence matrix
```{r}
csi_pres_abs <- (otu_final > 0) * 1
```

create relative abundance matrices
```{r}
csi_relabun <- otu_final
for (i in 1:dim(otu_final)[1]) {
csi_relabun[i, ] <- otu_final[i, ]/sum(otu_final[i,
])
}
```

import taxonomy file
```{r}
csi_tax <- read.tax(taxonomy = "../data/CSI.0.03.cons.taxonomy", format = "rdp", tax.levels = 6, col.tax = 3)
```        



bind design and bact files
```{r}
new_csi <- cbind(design_final,csi_relabun)
df <- decostand(csi_relabun, method="standardize") 

help(decostand)
#double check if you need hellinger transformation of relative abundance data 
#view details: i need to specify the method
```

subset data so bact and nutrients match (only first and last dates)
do not need to do this for salinity - need this???

***ALEX NOTES: confused.. why do we do this step?***
```{r}
csi_dat <- na.omit(new_csi)
dim(new_csi)
df_bact <- csi_dat[,-c(1:15)]
dim(df_bact)
df_nutr <- csi_dat[,c(9:11)]
dim(df_nutr)
```

set treatments (salinity levels and dispersal levels)
```{r}
treatments1 <- as.factor(design_final$Salinity)
levels(treatments1) <- c("0","5","9","13")
treatments2 <- as.factor(design_final$Dispersal)
levels(treatments2) <- c("2","3")

date_1 <- as.factor(design_final$Date2)
```

How much variation in community composition is due to environmental factors?
Run PERMANOVA using adonis function
```{r}
adonis = adonis(new_csi[,-c(1:15)] ~ Date2*Dispersal*Salinity, method = "bray", data = new_csi, perm=1000)
adonis
```

import zooplankton data
```{r}
zoop <- read.csv("../data/zoop_CSI.csv", row.names=1)
str(zoop)
dim(zoop)
```

remove low samples
```{r}
#removed low samples
zoop.in <- zoop[,-c(1:6)]
dim(zoop.in)
```

remove sample CSI026 from zooplankton community data
```{r}
odd.sites <- c("CSI026")
zoop.in2 <- zoop.in[setdiff(rownames(zoop.in), odd.sites), ]
dim(zoop.in2)
```

remove sample CSI026 from bacterial community data
```{r}
otu_final.2 <- otu_final[setdiff(rownames(otu_final), odd.sites), ]
dim(otu_final.2)
```

make relative abundance matrix without CSI026
```{r}
zoopREL <- zoop.in2
for(i in 1:dim(zoop.in2)[1]){
  zoopREL[i,] <- zoop.in2[i,]/sum(zoop.in2[i,])
}

csi_relabun.2 <- otu_final.2
for(i in 1:dim(otu_final.2)[1]){
  csi_relabun.2[i,] <- otu_final.2[i,]/sum(otu_final.2[i,])
}
```

Is there a relationship between zooplankton and bacterial community composition?

matrix comparison of zoop and bacterial communities using mantel test
```{r}
dist.zoop <- vegdist(zoopREL, method = "bray")
dist.bact <- vegdist(csi_relabun.2, method = "bray")

mantel.rtest(dist.zoop, dist.bact, nrepet = 999)
```

DIVERSITY METRICS

rarefy abundances
```{r}
rowSums(otu_final)
otu_rarefy <- rrarefy(otu_final, 13000)
```

for diversity metrics, use code from **CropDiversity repo
```{r}
# Fisher's Alpha
fisher <- fisher.alpha(otu_rarefy)
fisher

# Species Richness
richness <- rowSums((otu_rarefy >= 1))

# Shannon Diversity
shannon <- diversity(otu_rarefy, "shannon")

# Simpson's Evenness
simp.even <- apply(otu_rarefy, 1, simp_even)

#Pielou’s evenness
J <- shannon/log(specnumber(otu_rarefy[,-c(1:1)]))

#combined richness, diversity, evenness
CSI.DIVsoilmicrobes <- cbind(design_final,richness,shannon,simp.even,J)
write.table(CSI.DIVsoilmicrobes, file="../analyses/CSI.DIVsoilmicrobes.csv", sep=",", col.names=NA)
```

# Hypothesis Testing
```{r}
library("agricolae")
# First check the order
length(design_final$Salinity) == length(fisher)
all.equal(gsub("-", "", rownames(design_final)), names(fisher))

fisher.lm <- lm(fisher ~ Date2*Dispersal*Salinity, data = design_final)
anova(fisher.lm)
summary(fisher.lm)
HSD <- HSD.test(fisher.lm,"Salinity", console=TRUE)

richness.lm <- lm(richness ~ Date2*Dispersal*Salinity, data = design_final)
anova(richness.lm)
summary(richness.lm)
HSD <- HSD.test(richness.lm,"Salinity", console=TRUE)

shannon.lm <- lm(shannon ~ Date2*Dispersal*Salinity, data = design_final)
anova(shannon.lm)
summary(shannon.lm)
HSD <- HSD.test(shannon.lm,"Salinity", console=TRUE)

simp.even.lm <- lm(simp.even ~ Date2*Dispersal*Salinity, data = design_final)
anova(simp.even.lm)
summary(simp.even.lm)

J.lm <- lm(J ~ Date2*Dispersal*Salinity, data = design_final)
anova(J.lm)
summary(J.lm)
```

plot shannon diversity - but haven't fixed legend
```{r}
p <- ggplot(CSI.DIVsoilmicrobes, aes(x=Salinity, y=shannon, color=Dispersal))+geom_point()
p1=p+geom_smooth(method="lm",aes(fill=Dispersal))
p1 + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) +theme(axis.title=element_text(vjust=1,size=16,face="bold"), axis.text=element_text(size=14), axis.text.x = element_text(vjust=0.65, hjust=0.5, size=14), panel.border = element_rect(colour = "black",size=1.25)) + theme(axis.ticks.length=unit(0.3,"cm")) + xlab("Salinity") + ylab("Shannon Diversity Index") + scale_fill_discrete(name = "Dispersal",labels = c("Disp2", "Disp3"))

ggsave("../analyses/Cmin.pdf", plot=last_plot(), device=NULL, path=NULL, scale=1, width=NA, height=NA, dpi=300, limitsize=TRUE)
```

Alex extra code unused 
```{r}
bact_dist <- vegdist(csi_fin_rel, method="bray")
env_dist <- vegdist(scale(df_env), "euclid") #did not work - b/c there were NA's
nutr_dist <- vegdist(df_nutr)
```

alpha diversity (shannon weiner)

```{r}
a_div <- diversity(csi_fin)
a_div
```

beta diversity - confusing

"Function betadiver finds all indices reviewed by Koleff et al. (2003). All these indices could be found with function designdist which uses different notation, but the current function provides a conventional shortcut. The function only finds the indices. The proper analysis must be done with functions such as betadisper, adonis or mantel.""

```{r}
b_div <- betadiver(csi_fin, method = "w")
b_div
```

attempting diversity metrics (from stormwater) 
THINK ABOUT LOOKING AT DIV ACCORDING TO BOTH SALINITY AND DISPERSAL
HOW TO DO THIS?

```{r}
fisher <- fisher.alpha(otu_rarefy)
fisher

richness <- rowSums((otu_rarefy >= 1))
richness

shannon <- diversity(otu_rarefy, "shannon")
shannon

simp.even <- apply(otu_rarefy, 1, simp_even)
simp.even

colnames(design.r)

```

attempting diversity function plots (from stormwater)

```{r}
design_final
otu_final

otu_n_design <- cbind(design_final,otu_final)

design.r <- design_final[which(rownames(design_final) %in% rownames(otu_final)), ]

design.r

design.disp <- design.r[design.r$Dispersal != "W", ]
design.sal <- design.r[design.r$Salinity != "w", ]
design.date <- design.r[design.r$Date2 != "W", ]

disp.bac <- as.data.frame(matrix(NA, dim(design.disp)[1], 10))
sal.bac <- as.data.frame(matrix(NA, dim(design.sal)[1], 10))
date.bac <- as.data.frame(matrix(NA, dim(design.date)[1], 10))

colnames(disp.bac) <- c("Field_ID","Replicate","Date2","Treatment","Salinity","Dispersal","S","F_Alpha","S_Even","Shannon")

disp.bac$Sample <- rownames(design.disp)
disp.bac$Field_ID <- design.disp$Field_ID
disp.bac$Replicate <- design.disp$Replicate
disp.bac$Date2 <- design.disp$Date2
disp.bac$Treatment <- design.disp$Treatment
disp.bac$Salinity <- design.disp$Salinity
for(i in 1:length(rownames(design.disp))){
  temp <- rownames(design.disp)[i]
  disp.bac$S[i] <- richness[which(names(richness) == temp)]
  disp.bac$F_Alpha[i] <- fisher[which(names(fisher) == temp)]
  disp.bac$S_Even[i] <- simp.even[which(names(simp.even) == temp)]
  disp.bac$Shannon[i] <- shannon[which(names(shannon) == temp)]
}

disp.bac.m <- melt(disp.bac, id = c("Sample","Date2","Replicate","Salinity"))
disp.bac.c <- cast(data = disp.bac.m, Sample + Date2 ~ variable, c(mean,se), na.rm=T)

disp.bac.c <- as.data.frame(disp.bac.c)
disp.bac.m <- as.data.frame(disp.bac.m)

png(filename="../analyses/TEST_Div_Fxn.png",
    width = 1200, height = 800, res = 96*2)

test_plot <- barplot(disp.bac.c[, 5],
                     ylab="",xlab="",
                     lwd=3,col="gray",cex.lab=1.25,cex.names = 1.25, 
                     ylim = c(0,4500),yaxt="n",
                     space=c(1,0.25,1,0.25,1,0.25),
                     density=c(-1,15,-1,15,-1,15))
arrows(x0=test_plot, y0=disp.bac.c[, 5], y1 = disp.bac.c[, 5] - disp.bac.c[, 6], angle=90, length=0.1, lwd=2)
arrows(x0=test_plot, y0=disp.bac.c[, 5], y1 = disp.bac.c[, 5] + disp.bac.c[, 6], angle=90, length=0.1, lwd=2)
axis(side=2, labels=T,lwd.ticks=2,las=2,lwd=2,at=c(seq(0,150,75)))
abline(h=0,lwd=2,lty=3)

#need to figure this out -- how to make diversity figures? very different from stormwater example... maybe need another method
```

CHANGING PACE - RELATIONSHIP B/T COMM COMP AND DECOMP RATES
dbRDA and Mantel

Mantel - similarity between crobe and zoop communities? (based off mantel test in IL Wetlands looking at crobe/plant)

```{r}
zoop_csi <- read.csv("../data/zoop_CSI.csv")

rownames(zoop_csi) <- rownames(design_final)

all.equal(rownames(zoop_csi),rownames(design_final))

design_final

zoop_csi
#which relabun matrix to use? may have to re-create
```

